Instructions for ACL 2019 Proceedings

### 2023-12-10 paper/paper.tex
update tex
panic model adjustments, paper additions

@@

Prior research has attempted to create neural networks to function as
language acquisition devices (LAD) in order to model how children acquire
the English past tense. Most notably, Rumelhart and McClelland (1986)
made use of a perceptron model to model the learning of English past
tense in three stages based on prior research into how children
acquire it, and obtained learning results that very closely
align with children's learning. We therefore investigate whether
the acquisition of the genitive case in Russian would yield similar results.

@@

Russian, being a member of the Slavic branch of the Indo-European languages,
has a fully functioning grammatical case system which is used for all nouns
unlike English. Nouns and noun modifiers in Russian inflect for case
by changing the ending of the word, and the exact manner by which an ending
changes depends on the ending, case, and number of the noun. Table 1 shows 
the different possible endings of the Russian nominative and genitive cases in 
the singular and plural.

@@

Rumelhart and McClelland (1986) made use of a scheme to encode
English phoneme units, known as Wickelphones and Wickelfeatures, described
in Wickelgren (1969). Wickelphones are phoneme units representing
each phone in a word as triples, which consist of the phoneme itself,
its predecessor to the left, and its successor on the right.
Wickelfeatures are a means of representing Wickelphones as distributed
patterns of activation by capturing one feature of the central phoneme,
one feature of the predecessor, and one feature of the successor.

@@

Table 2 shows the scheme that we used to categorize Russian phonemes
on five simple dimensions.

\begin{table*}[t!]
\centering
\begin{tabular}{cccccccccccccc}
    \multicolumn{2}{c}{} & \multicolumn{12}{c}{Place}\\\hline
    \multicolumn{2}{c}{} & \multicolumn{4}{c}{Front} & \multicolumn{4}{c}{Middle} & \multicolumn{4}{c}{Back}\\\hline
    \multicolumn{2}{c}{} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced}\\\hline
    \multicolumn{2}{c}{} & H & S & H & S & H & S & H & S & H & S & H & S\\\hline
    \multirow{2}{50px}{Interrupted} & Stop & p & p' & b & b' & t & t' & d & d' & k & - & g & -\\
    & Nasal & - & - & m & m' & - & - & n & n' & - & - & - & - \\
    \multirow{2}{50px}{Cont.Cons.} & Fricative & f & f' & v & v' & s & s' & z & z' & s\` & t\_s\textbackslash & z\` & -\\
                                        & Liq/Sem. & - & - & l & l' & t\_s & - & r & r' & x & - & - & j\\
    \multirow{2}{50px}{Vowel} & Close & - & - & i & - & - & - & - & - & - & - & u & -\\
                              & Open & - & - & e & - & - & - & a & - & - & - & o & -\\
\end{tabular}
\caption{Categorization of Russian phonemes on five dimensions.
All phonemes are written using X-SAMPA.}
\end{table*}

@@

% TODO: explain this better
After being converted into X-SAMPA, each phoneme would then be converted into
its Wickelphone/Wickelfeature representations.

The Wickelfeatures would then be converted into arrays of activations for
certain features.

@@

\textbf{Citations}: Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text
itself, as Gusfield~\shortcite{Gusfield:97}.
Using the provided \LaTeX\ style, the former is accomplished using
{\small\verb|\cite|} and the latter with {\small\verb|\shortcite|} or {\small\verb|\newcite|}. Collapse multiple citations as in~\cite{Gusfield:97,Aho:72}; this is accomplished with the provided style using commas within the {\small\verb|\cite|} command, \emph{e.g.}, {\small\verb|\cite{Gusfield:97,Aho:72}|}. Append lowercase letters to the year in cases of ambiguities.  
 Treat double authors as
in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
as sentence constituents.

\section*{Acknowledgments}

The acknowledgments should go immediately before the references.  Do
not number the acknowledgments section. Do not include this section
when submitting your paper for review. \\

\noindent \textbf{Preparing References:} \\
Include your own bib file like this:
\verb|\bibliographystyle{acl_natbib}|
\verb|\bibliography{acl2019}|

where \verb|acl2019| corresponds to a acl2019.bib file.
\bibliography{acl2019}
\bibliographystyle{acl_natbib}

>>>>

\newcommand{\citenp}[1]{\citeauthor{#1} (\citeyear{#1})}

@@

Prior research has attempted to create neural networks to function as
language acquisition devices (LAD) in order to model how children acquire
the English past tense. Most notably, \citenp{RMcC}
made use of a perceptron model to model the learning of English past
tense in three stages based on prior research into how children
acquire it, and obtained learning results that very closely
align with children's learning. We therefore investigate whether
the acquisition of the genitive case in Russian would yield similar results.

@@

Russian, being a member of the Slavic branch of the Indo-European languages,
has a fully functioning grammatical case system which is used for all nouns
unlike English. Nouns and noun modifiers in Russian inflect for case
by changing the ending of the word, and the exact manner by which an ending
changes depends on the ending, case, and number of the noun. Table \ref{declension} shows 
the different possible endings of the Russian nominative and genitive cases in 
the singular and plural.

@@

\citenp{RMcC} made use of a scheme to encode
English phoneme units, known as Wickelphones and Wickelfeatures, described
in \citenp{Wickelgren}. Wickelphones are phoneme units representing
each phone in a word as triples, which consist of the phoneme itself,
its predecessor to the left, and its successor on the right.
Wickelfeatures are a means of representing Wickelphones as distributed
patterns of activation by capturing one feature of the central phoneme,
one feature of the predecessor, and one feature of the successor.

@@

Table \ref{phonemes} shows the scheme that we used to categorize Russian phonemes
on five simple dimensions.

\begin{table*}[t!]
\centering
\begin{tabular}{cccccccccccccc}
    \multicolumn{2}{c}{} & \multicolumn{12}{c}{Place}\\\hline
    \multicolumn{2}{c}{} & \multicolumn{4}{c}{Front} & \multicolumn{4}{c}{Middle} & \multicolumn{4}{c}{Back}\\\hline
    \multicolumn{2}{c}{} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced} & \multicolumn{2}{c}{Voiceless} & \multicolumn{2}{c}{Voiced}\\\hline
    \multicolumn{2}{c}{} & H & S & H & S & H & S & H & S & H & S & H & S\\\hline
    \multirow{2}{50px}{Interrupted} & Stop & p & p' & b & b' & t & t' & d & d' & k & - & g & -\\
    & Nasal & - & - & m & m' & - & - & n & n' & - & - & - & - \\
    \multirow{2}{50px}{Cont.Cons.} & Fricative & f & f' & v & v' & s & s' & z & z' & s\` & t\_s\textbackslash & z\` & -\\
                                        & Liq/Sem. & - & - & l & l' & t\_s & - & r & r' & x & - & - & j\\
    \multirow{2}{50px}{Vowel} & Close & - & - & i & - & - & - & - & - & - & - & u & -\\
                              & Open & - & - & e & - & - & - & a & - & - & - & o & -\\
\end{tabular}
\caption{\label{phonemes} Categorization of Russian phonemes on five dimensions.
All phonemes are written using X-SAMPA.}
\end{table*}

@@

However, phonemic representations are not specific enough for the neural network.
As previously referenced, \citenp{RMcC} refer to \citenp{Wickelgren}
to devise a featurization system using Wickelphones and Wickelfeatures.
In particular, each Wickelphone has a set of at most 25 Wickelfeatures that its presence activates,
which are derived from the phonological features of all three phonemes present in the Wickelphone.

For all phonemes other than the special word-boundary marker (which has a single unique `boundary' feature),
each phoneme has five features that correspond to the dimensions outlined in Table \ref{phonemes}.
For each feature in the central phoneme of the Wickelphone, at most five Wickelfeatures are generated:
each feature in the left context phoneme, the selected feature of the central phoneme,
and each feature in the right context. An example of this for the Wickelphone
\textbf{\textsubscript{j}a\textsubscript{z}} is present in Table \ref{features}.
Computationally, this algorithm corresponds well to a series of vector repetitions and a matrix transposition.

Each of these activated Wickelfeatures are then ``blurred": for a Wickelfeature $\left< f_1, f_2, f_3 \right>$,
all Wickelfeatures of the form $\left< * , f_2, f_3 \right>$ and $\left< f_1, f_2, * \right>$ have a random chance of being activated
as well. \citenp{RMcC} claim that this improves the neural network's ability to generalize between
phonological features.

However, to avoid computational blowup, Wickelfeatures with different dimensions for their left and right features
(other than the special word boundary feature) are filtered out of the computation. This brings the total number
of possible Wickelfeatures down from 2028 to only 660.

Activations across Wickelfeatures are thus represented by a boolean 660-vector, where a 1 in index $k$ corresponds to
Wickelfeature $k$ being activated. A word's total activation is all the Wickelfeatures activated by the word's composite Wickelphones. This is easily implemented with a vector sum operation.

\begin{table}[t!]
\begin{center}
	\small
\begin{tabular}{r|c|c|c}
	\# & Left Context & Central Phoneme & Right Context \\ \hline
	1 & Continuous & Vowel & Continuous \\
	2 & Liquid & Vowel & Fricative \\
	3 & Back & Vowel & Middle \\
	4 & Voiced & Vowel & Voiced \\
	5 & Soft & Vowel & Hard \\
	6 & Continuous & Open & Continuous \\
	7 & Liquid & Open & Fricative \\
	8 & Back & Open & Middle \\
	9 & Voiced & Open & Voiced \\
	10 & Soft & Open & Hard \\
	11 & Continuous & Middle & Continuous \\
	12 & Liquid & Middle & Fricative \\
	13 & Back & Middle & Middle \\
	14 & Voiced & Middle & Voiced \\
	15 & Soft & Middle & Hard \\
	16 & Continuous & Voiced & Continuous \\
	17 & Liquid & Voiced & Fricative \\
	18 & Back & Voiced & Middle \\
	19 & Voiced & Voiced & Voiced \\
	20 & Soft & Voiced & Hard \\
	21 & Continuous & Hard & Continuous \\
	22 & Liquid & Hard & Fricative \\
	23 & Back & Hard & Middle \\
	24 & Voiced & Hard & Voiced \\
	25 & Soft & Hard & Hard \\
\end{tabular}
\end{center}
\caption{\label{features} The Wickelfeatures activated by the Wickelphone
		                  \textbf{\textsubscript{j}a\textsubscript{z}}.}
\end{table}

@@

During training, the network's input is the activation vector for the nominative singular form
of a noun, and the expected output is the activation vector for that noun's genitive form
(singular or plural, depending on the network being trained). However, the larger model attempting
to learn both cases at once simply has the singular and plural genitive activation vectors concatenated together,
with the intent being that the model will eventually output two ``words" at once.

\subsection{Decoding model output}
Described in the Appendix of their paper, \citenp{RMcC} use what they term a ``binding network" to decode the feature activation vectors
outputted by the learning network into a full phonological representation. Because the activation
vector encoding scheme has no representation of the temporal dimension of a word (i.e., it does
not directly encode \textit{where} features and source phonemes occur in the source word), a separate
decoder is necessary. The binding network does not learn its weights, but rather each weight is determined
by the strength of its previous output ``in time", proportional to the probability of the corresponding feature
being activated (i.e., the network output).

The network attempts to model how each Wickelphone contributes to the Wickelfeatures present in the activation
vector, with the idea being ``to find a set of output features [Wickelphones] that accounts for as many as possible of the output features while minimizing the number of input features [Wickelfeatures] accounted for by more than one output feature. Thus, we want each of the output features to \textit{compete} for input features" \cite[pp. 269]{RMcC}.

\section{Results}
Unfortunately, our results were decidedly inconclusive.
The neural networks did successfully learn during training,
with a final loss of approximately 0.1 (down from 0.16 after 100 epochs).
However, the decoder network did not work as hoped, and only output two Wickelphones
when given output from any of the neural networks. For example, when given the network predicted activations
for the input \foreignlanguage{russian}{язык} (X-SAMPA \textbf{jazik}), the decoder only ever output
the Wickelphones \textbf{\textsubscript{\#}j\textsubscript{a}} and \textbf{\textsubscript{j}a\textsubscript{\#}};
for the input \foreignlanguage{russian}{жена} (X-SAMPA \textbf{z'ena}), the decoder only returned \textbf{\textsubscript{\#}z'\textsubscript{e}} and \textbf{\textsubscript{z'}e\textsubscript{\#}}.

\subsection{Discussion}
These results seem to suggest that a featural representation of words that properly preserves temporal
information is necessary to get meaningful output from these models. It is also possible that the
specification for the binding network in \citenp{RMcC} was overly general, and thus some nuance in implementation
may have been missed.

\section{Conclusion}

\bibliography{paper}{}
\bibliographystyle{acl_natbib}
